### 2013-09-30 tex/scam/appendix.tex
various edits and typo fixes

Let $\hat{r} \coloneqq Y - \bar{Y} \mathbf{1}_n - \sum_{k \in S} (\Delta_k \hat{d}_k - \hat{c}_k \mathbf{1}_n) $ be the residue.

@@

\begin{align*}
\partial d_k :& \qquad -\frac{1}{n} \Delta_k^\tran ( r_k - \Delta_k \d_k 
       + \c_k \mathbf{1}) + \lambda \mathbf{u}_k  - \mu_k - \gamma_k \Delta_k^\tran \mathbf{1} \\
\partial c_k: & \qquad -\frac{1}{n} \mathbf{1}^\tran ( \r_k - \Delta_k \d_k 
       + \c_k \mathbf{1}) + \gamma_k
\end{align*}
In the derivatives, $\mathbf{u}$ is a $(n-1)$-vector whose first coordinate is $\partial | d_{k1}|$ and all other coordinates are 1.

@@

Suppose $\lambda_n \geq c b (sB + \sigma) \sqrt{ \frac{s}{n} \log n \log (pn)}$, then with probability at least $ 1 - \frac{C}{n}$, for all $j,k$,
\[
\lambda_n >  | \frac{1}{n} \hat{r}^\tran \Delta_{k,j}|
\]
And therefore, the solution to the optimization~\ref{opt:alternate_opt} is zero on $S^c$.
\end{theorem}

\begin{proof}
The key is to note that $\hat{r}$ and $\Delta_{k,j}$ are independent for all $k,j$ because $\hat{r}$ is only dependent on $X_{S}$.

@@

\begin{align*}
\hat{r}_i &= Y_i - \sum_{k \in S} \hat{f}_k (X_k^{(i)}) \\
	&= f^*(X_S^{(i)}) + \epsilon_i - \bar{f^*} - \bar{\epsilon}- \sum_{k \in S} \hat{f}_k(X_k^{(i)}) \\
	&=  f^*(X_S^{(i)}) - \bar{f^*} - \sum_{k \in S} \hat{f}_k(X_k^{(i)}) + \epsilon_i - \bar{\epsilon}
\end{align*}
Where $\bar{f^*} = \frac{1}{n} \sum_{i=1}^n f^*(X_S^{(i)})$ and likewise for $\bar{\epsilon}$.

@@

We know that $|f^*(x)| \leq B$ and $|\hat{f}_k(x_k)| \leq B$ for all $k$.

Then $|\bar{f^*}| \leq B$ as well, and $|f^*(X_S^{(i)}) - \bar{f^*} - \sum_{k \in S} \hat{f}_k(X_k^{(i)})| \leq 3s B$.

@@

\textbf{Step 2.} We now bound $\frac{1}{n} \hat{r}^\tran \max( X, X_{(j)} \mathbf{1})$.

\[
\frac{1}{n} \hat{r}^\tran \max(X, X_{(j)} \mathbf{1}) = \frac{1}{n} \sum_{i=1}^n \hat{r}_i \max(X_i, X_{(j)}) = \frac{1}{n} \sum_{i=1}^n \hat{r}_i X_i \delta(ord(i) \geq j) + \frac{1}{n} X_{(j)} \mathbf{1}_A^\tran \hat{r}_A
\]

@@

\textbf{Term 1.}
\[
\trm{Want to bound} \qquad F(X_1,...,X_n) \coloneqq \frac{1}{n}\sum_{i=1}^n \hat{r}_i X_i \delta(ord(i) \geq j) 
\]

First, we note that $X_i$ is bounded in the range $[-b,b]$.

We claim then that $F$ is coordinatewise-Lipschitz. Let $X = (X_1, X_2,...,X_n)$ and $X' = (X'_1, X_2, ..., X_n)$ differ only on the first coordinate.

The order of coordinate $i$ in $X$ and $X'$ can change by at most $1$ for $i \neq 1$. Therefore, of the $j-1$ terms of the series, at most 2 terms differ from $F(X)$ to $F(X')$. Therefore,

$$ | F(X_1, ..., X_n) - F(X'_1, ..., X_n) | \leq \frac{4 b \|\hat{r}\|_\infty}{n} $$

By McDiarmid's inequality therefore, 
\[
P( | F(X) - \E F(X) | \geq t ) \leq C \exp( - c n\frac{t^2}{ (4 b \|\hat{r}\|_\infty)^2})
\]

By symmetry and the fact that $\hat{r}$ is centered, $\E F(X) = 0$.

We can fold the 4 into the constant $c$. With probability $1-\delta$, $|F(X)| \leq b \|\hat{r}\|_\infty \sqrt{   \frac{1}{cn} \log \frac{C}{\delta}}$.

@@

\[
\trm{Want to bound}   \frac{1}{n} X_{(j)} \mathbf{1}_A^\tran \hat{r}_A
\]

@@

Since $|X_{(j)}|$ is at most $b$, we obtain that with probability at least $1-\delta$, $| \frac{1}{n} X_{(j)} \mathbf{1}_A^\tran \hat{r}_A| \leq b\| \hat{r} \|_\infty \sqrt{ \frac{1}{cn} \log \frac{C}{\delta}}$.

@@

\[
| \frac{1}{n} \max(X, X_{(j)}\mathbf{1})^\tran \hat{r}| \leq b \|\hat{r}\|_\infty \sqrt{ \frac{1}{c} \frac{1}{n} \log \frac{np C }{\delta}}
\]

@@

$| \frac{1}{n} \max(X, X_{(j)}\mathbf{1})^\tran \hat{r}|$ is at most 
\[
 c b (sB + \sigma) \sqrt{ \frac{s}{n} \log n \log (pn) }
\]

>>>>

Let $\hat{r} \coloneqq Y - \bar{Y} \mathbf{1}_n - \sum_{k \in S} (\Delta_k \hat{d}_k - \hat{c}_k \mathbf{1}_n) $ be the residual.

@@

\begin{align*}
\partial d_k :& \qquad -\frac{1}{n} \Delta_k^\tran ( r_k - \Delta_k d_k 
       + c_k \mathbf{1}) + \lambda \mathbf{u}_k  - \mu_k - \gamma_k \Delta_k^\tran \mathbf{1} \\
\partial c_k: & \qquad -\frac{1}{n} \mathbf{1}^\tran ( r_k - \Delta_k d_k 
       + c_k \mathbf{1}) + \gamma_k
\end{align*}
In the derivatives, $\mathbf{u}$ is a $(n-1)$-vector whose first coordinate is $\partial | d_{k1}|$ and all other coordinates are 1.

@@

Suppose $\lambda_n \geq c b (sB + \sigma) \sqrt{ \frac{s}{n} \log n \log (pn)}$, then with probability at least $ 1 - \frac{C}{n}$, for all $j=1,...,n,k \in S^c$,
\[
\lambda_n >  | \frac{1}{n} \hat{r}^\tran \Delta_{k,j}|
\]
And therefore, the solution to the optimization~\ref{opt:alternate_opt} is zero on $S^c$.
\end{theorem}

\begin{proof}
The key is to note that $\hat{r}$ and $\Delta_{k,j}$ are independent for all $k \in S^c,j=1,...,n$ because $\hat{r}$ is only dependent on $X_{S}$.

We remind the reader that $\Delta_{k,j} = \max(X_k, - X_{k(j)} \mathbf{1}_n, 0)$. Because $\hat{r}$ is empirically centered,

\begin{align*}
\frac{1}{n} \hat{r}^\tran \Delta_{k,j} &= \frac{1}{n} \hat{r}^\tran \max(X_k, X_{k(j)}) 
                                            - \frac{1}{n} \hat{r}^\tran X_{k(j)} \\
                  &= \frac{1}{n} \hat{r}^\tran \max(X_k, X_{k(j)})
\end{align*}
Our goal in this proof is to bound $\frac{1}{n} \hat{r}^\tran \max(X_k, X_{k(j)})$ from above.

@@

\begin{align*}
\hat{r}_i &= Y_i - \bar{Y}_i -  \sum_{k \in S} \hat{f}_k (X_k^{(i)}) \\
	&= f_0(X_S^{(i)}) + \epsilon_i - \bar{f_0} - \bar{\epsilon}- \sum_{k \in S} \hat{f}_k(X_k^{(i)}) \\
	&=  f_0(X_S^{(i)}) - \bar{f_0} - \sum_{k \in S} \hat{f}_k(X_k^{(i)}) + \epsilon_i - \bar{\epsilon}
\end{align*}
Where $\bar{f_0} = \frac{1}{n} \sum_{i=1}^n f_0(X_S^{(i)})$ and likewise for $\bar{\epsilon}$.

@@

We know that $|f_0(x)| \leq sB$ and $|\hat{f}_k(x_k)| \leq B$ for all $k$.

Then $|\bar{f_0(x)}| \leq sB$ as well, and $|f^*(X_S^{(i)}) - \bar{f^*} - \sum_{k \in S} \hat{f}_k(X_k^{(i)})| \leq 3s B$.

@@

\textbf{Step 2.} We now bound $\frac{1}{n} \hat{r}^\tran \max( X, X_{k(j)} \mathbf{1})$.

\[
\frac{1}{n} \hat{r}^\tran \max(X_k, X_{k(j)} \mathbf{1}) = \frac{1}{n} \sum_{i=1}^n \hat{r}_i \max(X_{ki}, X_{k(j)}) = \frac{1}{n} \sum_{i=1}^n \hat{r}_i X_{ki} \delta(ord(i) \geq j) + \frac{1}{n} X_{k(j)} \mathbf{1}_A^\tran \hat{r}_A
\]

@@

\textbf{Term 1.}
\[
\trm{Want to bound} \qquad F(X_{k1},...,X_{kn}) \coloneqq \frac{1}{n}\sum_{i=1}^n \hat{r}_i X_{ki} \delta(ord(i) \geq j) 
\]

First, we note that $X_{ki}$ is bounded in the range $[-b,b]$.

We claim then that $F$ is coordinatewise-Lipschitz. Let $X_k = (X_{k1}, X_{k2},...,X_{kn})$ and $X'_k = (X'_{k1}, X_{k2}, ..., X_{kn})$ differ only on the first coordinate.

The order of coordinate $i$ in $X_k$ and $X'_k$ can change by at most $1$ for $i \neq 1$. Therefore, of the $j-1$ terms of the series, at most 2 terms differ from $F(X_k)$ to $F(X'_k)$ and

$$ | F(X_{k1}, ..., X_{kn}) - F(X'_{k1}, ..., X'_{kn}) | \leq \frac{4 b \|\hat{r}\|_\infty}{n} $$

By McDiarmid's inequality therefore, 
\[
P( | F(X_k) - \E F(X_k) | \geq t ) \leq C \exp( - c n\frac{t^2}{ (4 b \|\hat{r}\|_\infty)^2})
\]

By symmetry and the fact that $\hat{r}$ is centered, $\E F(X_k) = 0$.

We can fold the 4 into the constant $c$. With probability $1-\delta$, $|F(X_k)| \leq b \|\hat{r}\|_\infty \sqrt{   \frac{1}{cn} \log \frac{C}{\delta}}$.

@@

\[
\trm{Want to bound}   \frac{1}{n} X_{k(j)} \mathbf{1}_A^\tran \hat{r}_A
\]

@@

Since $|X_{k(j)}|$ is at most $b$, we obtain that with probability at least $1-\delta$, $| \frac{1}{n} X_{k(j)} \mathbf{1}_A^\tran \hat{r}_A| \leq b\| \hat{r} \|_\infty \sqrt{ \frac{1}{cn} \log \frac{C}{\delta}}$.

@@

\[
| \frac{1}{n} \max(X_k, X_{k(j)}\mathbf{1})^\tran \hat{r}| \leq b \|\hat{r}\|_\infty \sqrt{ \frac{1}{c} \frac{1}{n} \log \frac{np C }{\delta}}
\]

@@

$| \frac{1}{n} \max(X_k, X_{k(j)}\mathbf{1})^\tran \hat{r}|$ is at most 
\[
 c b (sB + \sigma) \sqrt{ \frac{s}{n} \log n \log (pn) }
\]

### 2013-09-30 tex/scam/sparsistency.tex
various edits and typo fixes

We follow Wainwright \cite{Wain:09a} and define a purely theoretical construct called \emph{restricted regression}.
\begin{definition}
In \emph{restricted regression}, we restrict the indices $k$ in
optimization (\ref{opt:alternate_opt}) to lie in the support $S$ instead of ranging from $1,...,p$. 
\end{definition}

Our analysis then diverges from the now-standard ``primal-dual witness
technique''~\cite{Wain:09a}. Primal-dual witness explicitly solves all the dual variables, but because our optimization is more complex, we do not solve the dual variables on $S$; we instead write the dual variables on $S^c$ as a function of the restricted regression \emph{residual}, which is implicitly a function of the dual variables on $S$.

\begin{theorem} (Deterministic setting)
\label{thm:deterministic}
Let $\{\hat{d}_k, \hat{c}_k\}_{k \in S}$ be the minimizer of the restricted regression, that is, the solution to optimization (\ref{opt:alternate_opt}) where we restrict $k \in S$. Let $\hat{d}_k = 0$ and $ \hat{c}_k = 0$ for $k \in S^c$.
Let $\hat{r} \coloneqq Y - \bar{Y}\mathbf{1}_n - \sum_{k \in S} (\Delta_k \hat{d}_k -
\hat{c}_k \mathbf{1}_n)$ be the restricted regression residual. For $k
\in \{1,...,p\}$, Let $\Delta_{k, j}$ be the $j$-th column of $\Delta_k$, i.e. the $n$-dimensional vector $\max( X_k - X_{k (j)} \mathbf{1}, 0)$. \\

@@

\begin{packed_enum}
\item Let $\mathcal{X} = [-b, b]^p$. Let $F$ be a distribution supported and positive on $\mathcal{X}$. Let $X^{(1)},..., X^{(n)} $ be iid samples from $F$.
\item Let $Y = f_0(X) + \epsilon$ where $f_0$ is the true function and $\epsilon$ is noise. Let $Y^{(1)},...,Y^{(n)}$ be iid samples.
\item Let $s \leq p$ be a positive integer. Let $S = \{1,...,s\}$ denote the relevant variables, i.e.,
  $f_0(X) = f_0(X_S)$.
\item Let $f_{01},...,f_{0s} \coloneqq \argmin_{f_1,...,f_s} \{ \E(f_0(X) - \sum_{k=1}^s f_k(X_k))^2 \,|\, \E[f_k(X_k)] = 0 \}$ where $\mu = \E f(X)$.
\end{packed_enum}

Each of our statistical theorems will use a subset of the following assumptions:
\begin{packed_enum}
\item[A1:] $X_S, X_{S^c}$ are independent.  \ A1': $\{ X_k \}_{k \in S}$ are independent.
\item[A2:] $\|f_0\|_\infty \leq sB$ \  A2': $f_0$ is convex,
  twice-differentiable, $L$-Lipschitz, and $\trm{supp}(f_0) = S$.
\item[A3:] Suppose $\epsilon$ is mean-zero sub-Gaussian, independent of $X$, with sub-Gaussian scale $\sigma$.
\item[A4:] For all $k=1,...,s$, $\E(f_{0k}(X_k))^2 \geq \alpha$ for some positive constant $\alpha$.
\end{packed_enum}

We will use assumptions A1, A2, A3 to control the probability of false positives and the stronger assumptions A1', A2', A3, A4 to control the probability of false negatives.

>>>>

We follow \cite{Wain:09a} and define the \emph{restricted regression} purely for theoretical purposes.
\begin{definition}
In \emph{restricted regression}, we restrict the indices $k$ in
optimization (\ref{opt:alternate_opt}) to lie in the support $S$ instead of ranging from $1,...,p$. 
\end{definition}

Our analysis then differs from the now-standard ``primal-dual witness
technique''~\cite{Wain:09a}. Primal-dual witness explicitly solves all the dual variables, but because our optimization is more complex, we do not solve the dual variables on $S$; we instead write the dual variables on $S^c$ as a function of the restricted regression \emph{residual}, which is implicitly a function of the dual variables on $S$.

\begin{theorem} (Deterministic setting)
\label{thm:deterministic}
Let $\{\hat{d}_k, \hat{c}_k\}_{k \in S}$ be the minimizer of the restricted regression, that is, the solution to optimization (\ref{opt:alternate_opt}) where we restrict $k \in S$. Let $\hat{d}_k = 0$ and $ \hat{c}_k = 0$ for $k \in S^c$.
Let $\hat{r} \coloneqq Y - \bar{Y}\mathbf{1}_n - \sum_{k \in S} (\Delta_k \hat{d}_k -
\hat{c}_k \mathbf{1}_n)$ be the restricted regression residual. For $k
\in \{1,...,p\}$, Let $\Delta_{k, j} \in R^n$ be the $j$-th column of $\Delta_k$, i.e. $\max( X_k - X_{k (j)} \mathbf{1}_n, 0)$. \\

@@

\begin{packed_enum}
\item Let $F$ be a distribution supported and positive on $\mathcal{X}=[-b,b]^p$. Let $X^{(1)},..., X^{(n)} \sim F$ be iid.
\item Let $Y = f_0(X) + \epsilon$ where $\epsilon$ is zero-mean noise. Let $Y^{(1)},...,Y^{(n)}$ be iid.
\item Let $S = \{1,...,s\}$ denote the relevant variables where $s\leq p$, i.e.,
  $f_0(X) = f_0(X_S)$.
\item Let $f_{01},...,f_{0s} \coloneqq \argmin_{f_1,...,f_s} \{ \E(f_0(X) - \sum_{k=1}^s f_k(X_k))^2 \,|\, \E[f_k(X_k)] = 0 \}$.
\end{packed_enum}

Each of our theorems will use a subset of the following assumptions:
\begin{packed_enum}
\item[A1:] $X_S, X_{S^c}$ are independent.  \ A1': $\{ X_k \}_{k \in S}$ are independent.
\item[A2:] $\|f_0\|_\infty \leq sB$ \  A2': $f_0$ is convex,
  twice-differentiable, $L$-Lipschitz, and $\trm{supp}(f_0) = S$.
\item[A3:] Suppose $\epsilon$ is mean-zero sub-Gaussian, independent of $X$, with sub-Gaussian scale $\sigma$.
\item[A4:] For all $k=1,...,s$, $\E(f_{0k}(X_k))^2 \geq \alpha$ for some positive constant $\alpha$.
\end{packed_enum}

We will use assumptions A1, A2, A3 to control the probability of false positives and the stronger assumptions A1', A2', A3, A4 to control the probability of false negatives. The $\alpha$ in assumption A4 can be interpreted as the signal strength of the relevant variable detection problem.
